---
title: "HW 10"
author: "Kate Spiekerman, kls5665"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r global_options, echo=FALSE}
knitr::opts_chunk$set(fig.height=4, fig.width=5, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60))
```

The github link can be found [here](https://github.com/kspiekerman/HW10).

```{r, echo=FALSE}
redline <- read.csv('redlining.csv')

library(ggplot2)
library(tidyverse)
```

## **Problem 1**

Linear Regression:

```{r, echo=FALSE}
model <- lm(policies ~ minority + fire + age + income, data = redline)
summary(model)

conf_int <- confint(model, level = 0.95)
```

1) Question: What question are you trying to answer?
We are trying to answer whether or not there is a linear correlation between the amount of FAIR policies issued and racial/ethnic patterns. Our goal is to determine if the variance in FAIR policies is due to reasonable, unbiased factors or ones relating to race and ethnicity of residents. 

2) Approach: What approach/statistical tool did you use to answer the question?
I used a multivariate linear regression to compare the number of FAIR policies in a certain zip code to the percentage of residents in a ZIP code who identified as a member of racial/ethnic minority. Additionally the model was adjusted to account for fires in the zip, percent of housing units in that ZIP code built before WWII, and median family income in thousands of dollars.  

3) Results: What evidence/results did your approach provide to answer the question? (E.g. any numbers,
tables, figures as appropriate.
Based on the data provided, the coefficient for minority is 0.008359, and the associated p-value is 0.00611. This is statistically significantly low. This indicates that there is positive linear correlation between the racial/ethnic composition and the number of FAIR policies issued.

4) Conclusion: What is your conclusion about your question? Provide a written interpretation of your
results, understandable to stakeholders who might plausibly take an interest in this data set.
The distribution of FAIR policies in zip codes was more directly associated with the racial/ethnic variable than the other three, indicating that there was likely discrimination among this data set. However, there is still a correlation between the other three variables and FAIR policies, so it is not certain that discrimination is at play in this situation. The 95% confidence interval of income includes 0, indicating that there is statistical uncertainty in whether or not higher income is associated with number of FAIR policies.


## **Problem 2**

Part A

```{r, echo=FALSE}
groceries <- read.csv('groceries.csv')

avg_prices <- groceries %>%
  group_by(Store) %>%
  summarize(avg_price = round(mean(Price), 2)) %>%
  arrange(desc(avg_price))

ggplot(avg_prices) +
  geom_col(aes(x = avg_price, y = Store), fill = 'skyblue') +
  labs(title = 'Average Prices by Store", x = "Average Price", y = "Store')

```


Part B

```{r, echo=FALSE}
store_counts <- groceries %>%
  group_by(Product) %>%
  summarise(num_stores = n_distinct(Store))

ggplot(store_counts, aes(x = num_stores, y = Product)) +
  geom_col(fill = 'skyblue') +
  labs(title = 'Number of Stores that Sell Each Product',
       x = 'Number of Stores',
       y = 'Product')

```


Part C

```{r, echo=FALSE}
model <- lm(Price ~ Product + Type, data = groceries)
summary(model)

```

The average price difference for the same product between convenience stores and ordinary grocery stores is between .09 and .66. Since this interval does not contain 0 it is statistically significant. This means that ordinary groceries stores charge statistically significantly less than convenience stores for the same products. 


Part D

```{r, echo=FALSE}
model_store <- lm(Price ~ Product + Store, data = groceries)
summary(model_store)

```

Based on the coefficients made from the regression model, the two stores that charge the lowest prices when comparing the same products are Kroger Fresh Fare and Walmart. These stores have the lowest coefficients associated with them, indicating lower prices compared to the other stores in the data set.

Based on the coefficients made from the regression model, the two stores that charge the highest prices when comparing the same products are Whole Foods and Wheatsville Food Co-Op. These stores have the highest coefficients associated with them, indicating higher prices compared to the other stores in the data set.


Part E

Based on the regression in Part D, the coefficients of Central Market and HEB are -.57 and -.65 respectively. As a result of the data I would claim Central Market charges a similar amount to HEB for the same product.  
This would mean that the higher prices at Central Market are a result of different products sold, not higher prices for the same products. 

Part F

```{r, echo=FALSE}
groceries <- groceries %>%
  mutate(Income10K = Income / 10000)

model_income <- lm(Price ~ Product + Income10K, data = groceries)
summary(model_income)


```

The coefficient for Income 10K is -0.014. This means that for each additional $10,000 increase in income, the price is expected to decrease by approximately 0.014 units, on average. However, the p-value associated with this is .14 which is too large to be considered statistically significant. Essentially, there is not enough evidence to conclude that consumers in poorer ZIP codes pay significantly more or less for the same product, on average, based on income alone. 





